---
title: "elena_skin_analysis"
author: "Peter Higgins"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(here)
library(janitor)
library(gt)
library(flextable)
library(gtsummary)
library(rstatix)
library(RColorBrewer)
library(reshape2)
library(ggthemes)
```

## Read in data

Start with FCP and CRP

```{r labs}
fcp <- readxl::read_excel(here("Oct FCP and CRP for charts.xlsx"), sheet = "FCP") %>% 
  group_by(mrn) %>% 
  arrange(mrn, desc(time_point)) %>% 
  mutate(delta_fcp = value - lag(value)) %>% 
  mutate(pct_change_fcp = 100*delta_fcp/lag(value)) %>% 
  ungroup() %>% 
  select(group:pct_change_fcp)

crp <- readxl::read_excel(here("Oct FCP and CRP for charts.xlsx"), sheet = "CRP") %>% 
  group_by(mrn) %>% 
  arrange(mrn, desc(time_point)) %>% 
  mutate(delta_crp = value - lag(value)) %>% 
  mutate(pct_change_crp = 100*delta_crp/lag(value)) %>% 
  ungroup() %>% 
  select(group:pct_change_crp)
```

Now read in Likert values

```{r likert}
path_likert <- readxl::read_excel(here("Oct Likert Numbers.xlsx"), sheet = "path") %>% 
  select(group:path_likert)

endo_likert <- readxl::read_excel(here("Oct Likert Numbers.xlsx"), sheet = "endo") %>% 
  select(group:endo_likert)

imaging_likert <- readxl::read_excel(here("Oct Likert Numbers.xlsx"), sheet = "imaging") %>% 
  select(group:imaging_likert)
```

Now read in demographics

```{r demog}
demog <- readxl::read_excel(here("Oct demographics organized.xlsx")) %>% clean_names() %>% 
  select(group_type:stelara_start) %>% 
  mutate(dob = as.Date(dob),
         stelara_start = as.Date(stelara_start)) %>% 
  rename(group = group_type)
```

## Demographic table with gtsummary and flextable

```{r demographics}
 demog %>% 
   select(-dob, -stelara_start) %>% 
   mutate(sex = case_when(sex == "F" ~ "Female",
                          sex =="M" ~ "Male")) %>% 
   mutate(tobacco = case_when(tobacco == "Y" ~ "Yes",
                          tobacco == "N" ~ "No")) %>% 
   mutate(race = case_when(race == "A" ~ "Asian or Pacific Islander",
                          race == "AA" ~ "African-American",
                          race == "C" ~ "Caucasian",
                          race == "HL" ~ "More than One Race",
                          race == "O" ~ "Other",
                          race == "U" ~ "Unknown")) %>% 
   tbl_summary(
    by = group, # split table by group
    missing = "no", # don't list missing data separately
    label = c(sex ~ "Sex", race ~ "Race",
              tobacco ~ "Tobacco Use",
              age_at_start ~ "Age"),
    type = c(sex, race, tobacco) ~ "categorical") %>%
  add_n() %>% # add column with total number of non-missing observations
  add_p() %>% # test for a difference between groups
  modify_header(label = "**Variable**") %>% # update the column header
  bold_labels() %>% 
   as_flex_table()

```

## Testing FCP and CRP

You can also embed plots, for example:

```{r fcp, echo=FALSE}
fcp %>% 
  filter(time_point == "AFTER") %>% 
  rstatix::t_test(value ~ group)

fcp %>% 
  filter(time_point == "BEFORE") %>% 
  rstatix::t_test(value ~ group)

fcp %>% 
  filter(time_point == "AFTER") %>% 
  rstatix::t_test(pct_change_fcp ~ group, detailed = TRUE) ->
fcp_t

fcp_t %>% 
  select(estimate1:estimate2, group1:conf.high) %>% 
  flextable()
```

The FCP in the NOSKIN group decreased by `r round(fcp_t$estimate1,2)` percent, while the calprotectin in the SKIN group decreased by `r round(fcp_t$estimate2,2)` percent, with a p value of `r round(fcp_t$p,3)` for this comparison.

```{r crp, echo=FALSE}
crp %>% 
  filter(time_point == "AFTER") %>% 
  rstatix::t_test(value ~ group)

crp %>% 
  filter(time_point == "BEFORE") %>% 
  rstatix::t_test(value ~ group)

crp %>% 
  filter(time_point == "AFTER") %>% 
  rstatix::t_test(pct_change_crp ~ group, detailed = TRUE) ->
crp_t

crp_t %>% 
  select(estimate1:estimate2, group1:conf.high) %>% 
  flextable()
```

The CRP in the NOSKIN group decreased by `r round(crp_t$estimate1,2)` percent, while the calprotectin in the SKIN group decreased by `r round(crp_t$estimate2,2)` percent, with a p value of `r round(crp_t$p,3)` for this comparison.

## Testing Likert for Path, Endo, Imaging

```{r likert_results}
path_likert %>% 
  rstatix::t_test(path_likert ~ group, detailed = TRUE)->
path_likert_result
path_likert_result %>% 
  select(estimate1:estimate2, group1:conf.high) %>% 
  flextable()

endo_likert %>% 
  rstatix::t_test(endo_likert ~ group, detailed = TRUE) %>% 
  select(estimate1:estimate2, group1:conf.high) %>% 
  flextable()

imaging_likert %>% 
  rstatix::t_test(imaging_likert ~ group, detailed = TRUE) %>% 
  select(estimate1:estimate2, group1:conf.high) %>% 
  flextable()
```

The pathology likert scale results demonstrated a difference of `r  path_likert_result$estimate` points higher for the SKIN group, with a confidence interval from `r  -path_likert_result$conf.low` to `r  -path_likert_result$conf.high`

## Plotting Likert Results

Use likert package examples

<https://github.com/jbryer/likert>

and example for ggplot here <https://jtr13.github.io/cc19/likert.html>

better explained here
http://rnotr.com/likert/ggplot/barometer/likert-plots/

and generally about likert plots here
http://daydreamingnumbers.com/blog/4-ways-to-visualize-likert-scales/ 

```{r path, echo=FALSE}
path_likert %>% 
  mutate(label = factor(case_when(
                  path_likert ==1 ~ "Significantly Worse",
                  path_likert ==2 ~ "Slightly Worse",
                  path_likert ==3 ~ "No Change",
                  path_likert ==4 ~ "Slightly Better",
                  path_likert ==5 ~ "Completely Better")))  %>% 
  group_by(group, label) %>% 
  add_count()  %>% 
  distinct(.keep_all = TRUE) %>% 
  select(-path_likert) %>% 
  pivot_wider(names_from = "label", values_from = "n") %>% 
  relocate(group, "Completely Better", "Slightly Better", group) %>%     replace_na(list(`Significantly Worse`=0)) ->
likert_path

d <- likert_path
d[,2:6] <- d[,2:6]/rowSums(d[,2:6])

mytitle <- "Pathology Likert Ratings"
mylevels <- c("Completely Better", "Slightly Better", "No Change", "Slightly Worse",  "Significantly Worse")

# Generate mid value of neutral category
numlevels <- length(d[1,])-1
numcenter <- ceiling(numlevels/2) + 1
d$midvalues <- d[,numcenter]/2
d_2<-cbind(d[,1],d[,2:ceiling(numlevels/2)], d$midvalues, d$midvalues,d[,numcenter:numlevels+1])
colnames(d_2)<-c("group",mylevels[1:floor(numlevels/2)],"Midlow",
                 "Midhigh",mylevels[numcenter:numlevels])

# Split into six categories (splitting neutral)
numlevels<-length(mylevels)+1
point1<-2
point2<-((numlevels)/2)+1
point3<-point2+1
point4<-numlevels+1

# Assign color to each category
numlevels<-length(d[1,])-1
temp.rows<-length(d_2[,1])
pal<-brewer.pal((numlevels-1),"RdBu")
pal[ceiling(numlevels/2)]<-"#DFDFDF"
legend.pal<-rev(pal)
pal<-c(pal[1:(ceiling(numlevels/2)-1)], pal[ceiling(numlevels/2)], 
       pal[ceiling(numlevels/2)], pal[(ceiling(numlevels/2)+1):(numlevels-1)])

# Generate new data frame including all information
d_3<-melt(d_2,id="group")
d_3$col<-rep(pal,each=temp.rows)
d_3$value<-d_3$value*100
d_3$group<-str_wrap(d_3$group, width = 40)
d_3$group<-factor(d_3$group, levels = d_2$group[order(-(d_2[,5]+d_2[,6]+d_2[,7]))])


d_3$col <- c("#0571B0", "#0571B0",
            "#92C5DE", "#92C5DE",
            "#DFDFDF", "#DFDFDF",
            "#DFDFDF", "#DFDFDF",
            "#F4A582", "#F4A582",
            "#CA0020", "#CA0020")

highs<-na.omit(d_3[(length(d_3[,1])/2)+1:length(d_3[,1]),])
lows<-na.omit(d_3[1:(length(d_3[,1])/2),])

ggplot() + geom_bar(data=highs, aes(x = group, y=value, 
                  fill=factor(col, 
                  levels = c("#CA0020", "#F4A582", "#DFDFDF"))),
                    position="stack", stat="identity", width = 0.5) +
  geom_bar(data=lows, aes(x = group, y=-value, fill=factor(col,
                  levels = c("#0571B0", "#92C5DE", "#DFDFDF"))),
           position="stack", stat="identity", width = 0.5) +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Percent", labels = mylevels, 
                      breaks=legend.pal, guide="legend") + 
  theme_fivethirtyeight() + 
  coord_flip() +
  labs(title=mytitle, y="",x="") +
  theme(plot.title = element_text(size=14, hjust=0.5)) +
  theme(axis.text.y = element_text(hjust=0)) +
  theme(legend.position = "bottom") 
```

```{r endo, echo=FALSE}
endo_likert %>% 
  mutate(label = factor(case_when(
                  endo_likert ==1 ~ "Significantly Worse",
                  endo_likert ==2 ~ "Slightly Worse",
                  endo_likert ==3 ~ "No Change",
                  endo_likert ==4 ~ "Slightly Better",
                  endo_likert ==5 ~ "Completely Better")))  %>% 
  group_by(group, label) %>% 
  add_count()  %>% 
  distinct(.keep_all = TRUE) %>% 
  select(-endo_likert) %>% 
  pivot_wider(names_from = "label", values_from = "n") %>% 
  relocate(group, "Completely Better", "Slightly Better", 
           "No Change") %>%    
  mutate(`Significantly Worse`=0) ->
likert_endo

d <- likert_endo
d[,2:6] <- d[,2:6]/rowSums(d[,2:6])

mytitle <- "Endoscopy Likert Ratings"
mylevels <- c("Completely Better", "Slightly Better", "No Change", "Slightly Worse",  "Significantly Worse")

# Generate mid value of neutral category
numlevels <- length(d[1,])-1
numcenter <- ceiling(numlevels/2) + 1
d$midvalues <- d[,numcenter]/2
d_2<-cbind(d[,1],d[,2:ceiling(numlevels/2)], d$midvalues, d$midvalues,d[,numcenter:numlevels+1])
colnames(d_2)<-c("group",mylevels[1:floor(numlevels/2)],"Midlow",
                 "Midhigh",mylevels[numcenter:numlevels])

# Split into six categories (splitting neutral)
numlevels<-length(mylevels)+1
point1<-2
point2<-((numlevels)/2)+1
point3<-point2+1
point4<-numlevels+1

# Assign color to each category
numlevels<-length(d[1,])-1
temp.rows<-length(d_2[,1])
pal<-brewer.pal((numlevels-1),"RdBu")
pal[ceiling(numlevels/2)]<-"#DFDFDF"
legend.pal<-rev(pal)
pal<-c(pal[1:(ceiling(numlevels/2)-1)], pal[ceiling(numlevels/2)], 
       pal[ceiling(numlevels/2)], pal[(ceiling(numlevels/2)+1):(numlevels-1)])

# Generate new data frame including all information
d_3<-melt(d_2,id="group")
d_3$col<-rep(pal,each=temp.rows)
d_3$value<-d_3$value*100
d_3$group<-str_wrap(d_3$group, width = 40)
d_3$group<-factor(d_3$group, levels = d_2$group[order(-(d_2[,5]+d_2[,6]+d_2[,7]))])


d_3$col <- c("#0571B0", "#0571B0",
            "#92C5DE", "#92C5DE",
            "#DFDFDF", "#DFDFDF",
            "#DFDFDF", "#DFDFDF",
            "#F4A582", "#F4A582",
            "#CA0020", "#CA0020")

highs<-na.omit(d_3[(length(d_3[,1])/2)+1:length(d_3[,1]),])
lows<-na.omit(d_3[1:(length(d_3[,1])/2),])

ggplot() + geom_bar(data=highs, aes(x = group, y=value, 
                  fill=factor(col, 
                  levels = c("#CA0020", "#F4A582", "#DFDFDF"))),
                    position="stack", stat="identity", width = 0.5) +
  geom_bar(data=lows, aes(x = group, y=-value, fill=factor(col,
                  levels = c("#0571B0", "#92C5DE", "#DFDFDF"))),
           position="stack", stat="identity", width = 0.5) +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Percent", labels = mylevels, 
                      breaks=legend.pal, guide="legend") + 
  theme_fivethirtyeight() + 
  coord_flip() +
  labs(title=mytitle, y="",x="") +
  theme(plot.title = element_text(size=14, hjust=0.5)) +
  theme(axis.text.y = element_text(hjust=0)) +
  theme(legend.position = "bottom")
```

```{r imaging, echo=FALSE}
imaging_likert %>% 
  mutate(label = factor(case_when(
                  imaging_likert ==1 ~ "Significantly Worse",
                  imaging_likert ==2 ~ "Slightly Worse",
                  imaging_likert ==3 ~ "No Change",
                  imaging_likert ==4 ~ "Slightly Better",
                  imaging_likert ==5 ~ "Completely Better")))  %>% 
  group_by(group, label) %>% 
  add_count()  %>% 
  distinct(.keep_all = TRUE) %>% 
  select(-imaging_likert) %>% 
  pivot_wider(names_from = "label", values_from = "n") %>% 
  relocate(group, "Completely Better", "Slightly Better", 
           "No Change") %>%    
  mutate(`Significantly Worse`=0) ->
likert_imaging

d <- likert_imaging
d[,2:6] <- d[,2:6]/rowSums(d[,2:6])

mytitle <- "Imaging Likert Ratings"
mylevels <- c("Completely Better", "Slightly Better", "No Change", "Slightly Worse",  "Significantly Worse")

# Generate mid value of neutral category
numlevels <- length(d[1,])-1
numcenter <- ceiling(numlevels/2) + 1
d$midvalues <- d[,numcenter]/2
d_2<-cbind(d[,1],d[,2:ceiling(numlevels/2)], d$midvalues, d$midvalues,d[,numcenter:numlevels+1])
colnames(d_2)<-c("group",mylevels[1:floor(numlevels/2)],"Midlow",
                 "Midhigh",mylevels[numcenter:numlevels])

# Split into six categories (splitting neutral)
numlevels<-length(mylevels)+1
point1<-2
point2<-((numlevels)/2)+1
point3<-point2+1
point4<-numlevels+1

# Assign color to each category
numlevels<-length(d[1,])-1
temp.rows<-length(d_2[,1])
pal<-brewer.pal((numlevels-1),"RdBu")
pal[ceiling(numlevels/2)]<-"#DFDFDF"
legend.pal<-rev(pal)
pal<-c(pal[1:(ceiling(numlevels/2)-1)], pal[ceiling(numlevels/2)], 
       pal[ceiling(numlevels/2)], pal[(ceiling(numlevels/2)+1):(numlevels-1)])

# Generate new data frame including all information
d_3<-melt(d_2,id="group")
d_3$col<-rep(pal,each=temp.rows)
d_3$value<-d_3$value*100
d_3$group<-str_wrap(d_3$group, width = 40)
d_3$group<-factor(d_3$group, levels = d_2$group[order(-(d_2[,5]+d_2[,6]+d_2[,7]))])


d_3$col <- c("#0571B0", "#0571B0",
            "#92C5DE", "#92C5DE",
            "#DFDFDF", "#DFDFDF",
            "#DFDFDF", "#DFDFDF",
            "#F4A582", "#F4A582",
            "#CA0020", "#CA0020")

highs<-na.omit(d_3[(length(d_3[,1])/2)+1:length(d_3[,1]),])
lows<-na.omit(d_3[1:(length(d_3[,1])/2),])

ggplot() + geom_bar(data=highs, aes(x = group, y=value, 
                  fill=factor(col, 
                  levels = c("#CA0020", "#F4A582", "#DFDFDF"))),
                    position="stack", stat="identity", width = 0.5) +
  geom_bar(data=lows, aes(x = group, y=-value, fill=factor(col,
                  levels = c("#0571B0", "#92C5DE", "#DFDFDF"))),
           position="stack", stat="identity", width = 0.5) +
  geom_hline(yintercept = 0, color =c("white")) +
  scale_fill_identity("Percent", labels = mylevels, 
                      breaks=legend.pal, guide="legend") + 
  theme_fivethirtyeight() + 
  coord_flip() +
  labs(title=mytitle, y="",x="") +
  theme(plot.title = element_text(size=14, hjust=0.5)) +
  theme(axis.text.y = element_text(hjust=0)) +
  theme(legend.position = "bottom")
```
